---
title: "Why don't limpets coexist in our model?"
output: pdf_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


I'm having trouble getting limpets to coexist in the model. I'll step through some of the troubleshooting I've done to try to figure out why.

```{r, echo=TRUE}
library(tidyverse)
source("foodwebs/intertidal_foodweb/intertidal_foodweb_model_modified.R")
source("foodwebs/intertidal_foodweb/do_intertidal_partition.R")
```

# The problem

When you run the simulation and look at the time series output, limpets seem to be fine

```{r}
read_csv("fd_results.csv") %>%
  gather(balanus_glandula:free_space, key = "species",
         value = "abundance") %>%
  ggplot(aes(x = time, y = abundance)) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~species, scales = "free") +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

\pagebreak

However, when you partition coexistence, their r-bar is negative, indicating that they do not coexist with the two barnacle species.

```{r}
read_csv("fd_part_3_total_unmodified.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

\pagebreak

# Barnacle survival rate and limpet coexistence

First I tried to decrease the survival rates of balanus glandula adults, because they are a top competitor, to see if limpets would coexist.

Decreased the survival rates of BG from 0.8 to 0.75 for both adults and recruits:

```{r}
read_csv("fd_part_3_total_lowerbarn.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

Did not help.

\pagebreak

Next, I decreased both barnacle species survival rates, down to 0.7.

```{r}
read_csv("fd_part_3_total_lowerbothbarn.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

Also did not help limpets, but changed barnacle coexistence dynamics quite a bit.

\pagebreak

Next, try balanus slightly higher than chthamalus, to reflect realistic dynamics.

```{r}
read_csv("fd_part_3_total_bghigher.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

Again did not help limpets. Indicates that barnacle survival rate isn't driving limpet coexistence dynamics.

\pagebreak

# Lower variance in larval supply and limpet coexistence

From FD paper: when variance in larval supply decreases, then limpet density increased. So the next scenario is to vary the variance in larval supply while the means are all high.

Across the top row: increasing variance in larval supply, with simulation run 3 times.

```{r}
read_csv("test_var_list_2_df.csv") %>%
  mutate(larval_scenario = factor(larval_scenario, levels = c("low", "med", "high"))) %>%
    mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = mean_cs)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  geom_errorbar(aes(ymin = mean_cs - se_cs, ymax = mean_cs + se_cs), color = "black", width = 0.2) +
  facet_grid(species ~ larval_scenario, scales = "free") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Mechanistic partitioning") +
  ylab("Growth rate when rare")
```

Still no positive coexistence from limpets.

\pagebreak

# Larval limpet limitation

Maybe limpet abundance is just too low - and ramping up larval supply could help them achieve dominance. So increased mean larval supply by 10-20x and ran for medium and high variance in supply (simulation failed at the "low" scenario).

```{r}
read_csv("test_var_list_3_df.csv") %>% 
  mutate(larval_scenario = factor(larval_scenario, levels = c("low", "med", "high"))) %>%
    mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = mean_cs)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  geom_errorbar(aes(ymin = mean_cs - se_cs, ymax = mean_cs + se_cs), color = "black", width = 0.2) +
  facet_grid(species ~ larval_scenario, scales = "free") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("Mechanistic partitioning") +
  ylab("Growth rate when rare")
```

Still no change in limpet coexistence.

\pagebreak

Because chthamalus is the subdominant barnacle competitor, maybe higher larval supply for CD could allow limpets to coexist? Try increasing CD larval supply:

```{r}
read_csv("fd_results_cd_df.csv") %>%
    mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = mean_cs)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  geom_errorbar(aes(ymin = mean_cs - se_cs, ymax = mean_cs + se_cs), color = "black", width = 0.2) +
  facet_wrap(~ species, scales = "free") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("Mechanistic partitioning") +
  ylab("Growth rate when rare")
```

Still doesn't help. 

\pagebreak

Try BOTH CD and limpets at high larval supply:

```{r}
read_csv("fd_results_cd_l_df.csv") %>%
    mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = mean_cs)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  geom_errorbar(aes(ymin = mean_cs - se_cs, ymax = mean_cs + se_cs), color = "black", width = 0.2) +
  facet_wrap(~ species, scales = "free") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("Mechanistic partitioning") +
  ylab("Growth rate when rare")
```

Nope.

\pagebreak

# Limpet "density-dependent" parameter

The limpet equation has a "density-dependent parameter" (delta) because there are no limpet predators in the model. It was set by Forde and Doak to  be -0.02, but for no real reason. I also have always had trouble getting limpet populations to stabilize close to their initial starting values, suggesting density dependence (or interspecific competition) is keeping it low.

Here, I lowered the density-dependent parameter by an order of magnitude and ran CD and limpet supply on high.

```{r}
read_csv("fd_results_cd_l_delta_df.csv") %>%
   mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = mean_cs)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  geom_errorbar(aes(ymin = mean_cs - se_cs, ymax = mean_cs + se_cs), color = "black", width = 0.2) +
  facet_wrap(~species, scales = "free") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("Mechanistic partitioning") +
  ylab("Growth rate when rare")
```

No improvement.

\pagebreak

# Limpet size and coexistence

Because competition here is determined based on a combination of abundance and individual size, perhaps limpets are not big enough in the model to coexist. 

Here, I changed the density dependent parameter back and updated limpet size and density based on [Gilman 2006](https://doi.org/10.1111/j.0906-7590.2006.04572.x).

```{r}
read_csv("fd_part_3_total_size.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

Also does not help coexistence. Might be able to come back to this and explore it more by pushing up limpet size and density, but don't have any biologically-motivated reason for doing so.

\pagebreak

# Simulation length

Perhaps simulation isn't going long enough? Run longer:

```{r}
read_csv("fd_part_3_total_years.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

\pagebreak

# Density-dependence and limpet recruitment

One problem seems to be that limpet recruitment is so severely density-dependent - it's saturating at a very low number in the model.

The number of new recruits added to adult population size is, if 700 recruits are in the larval pool and the population size is the starting population size:

```{r, echo = TRUE} 
0.88 * 700 * exp(-0.02 * 239)
```

Only 5 new adults! So if we remove the density dependent parameter all together:

```{r}
read_csv("fd_part_3_total_nod.csv") %>%
  gather(delta_0:delta_cp, key = "coexistence_partition", value = "coexistence_strength") %>%
  mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = coexistence_strength, color = coexistence_partition)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  facet_grid( ~ species) +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1))
```

That also does not help.

# Everything combined

Combine everything that might possibly help the limpets: decreased density dependence, larger size, higher density, and more larvae. Across the columns: larval supply levels for limpets alone were increased, starting at 1x supply to 10x supply. Nothing made a large difference.

```{r}
read_csv("test_var_list_all_df.csv") %>%
  mutate(larval_scenario = factor(larval_scenario, levels = c("low_x1", "med_x2", "high_x10"))) %>%
    mutate(coexistence_partition = factor(coexistence_partition, levels = c("r_bar", 
                                                                          "delta_0", "delta_c", 
                                                                          "delta_p", "delta_cp"))) %>%
  ggplot(aes(x = coexistence_partition, y = mean_cs)) +
  geom_bar(stat = "identity", aes(fill = coexistence_partition)) + 
  geom_errorbar(aes(ymin = mean_cs - se_cs, ymax = mean_cs + se_cs), color = "black", width = 0.2) +
  facet_grid(species ~ larval_scenario, scales = "free") +
  geom_hline(yintercept = 0) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90, hjust = 1)) + 
  xlab("Mechanistic partitioning") +
  ylab("Growth rate when rare")
```

# Summary

No matter what changes I make, I see the same generic pattern: no limpet coexistence, delta_c is very negative and delta_cp is very positive. This suggests to me there is something structural about the model that prevents the coexistence of limpets. They both do not have a positive invasion growth rate, are out competed by barnacles, and do not "benefit" from predation.

Honestly I'm not sure whether the limpets EVER coexisted in the original model.



