---
title: "Untitled"
output: github_document
---

## Niche and Cascade Models

There are a few functions already written to generate food webs using the cascade model and the niche model (Williams & Martinez 2000 Nature, doi: 10.1038/35004572), provided by Owen Petchy:

```{r}
source("petchey_food_web_functions.R")
```

For a given number of species (*S*), a given number of links (*L*), generate one (N = 1) or a list (N > 1) of networks as adjacency matrices.

```{r}
Cascade.model(S = 5, L = 5, N = 1)
```

```{r}
Niche.model(S = 5, L = 5, N = 1)
```

## Allometric Trophic Network (ATN) models

Following the example of an implementation by Iles & Novak (2016 The American Naturalist, doi: 10.1086/686730), the general steps for implementing an ATN model are:

1. Generate a food web using the niche or cascade model, for a chosen number of species and connectance  
2. Assign body sizes to the species, based on consumer-resource mass ratios  
3. Use as inputs to Yodzis & Innes's bioenergetic model of consumer-resource dynamics

This workflow has recently been implemented for **julia**, a different programming language, in Delmas et al. 2017 Methods in Ecology & Evolution (doi: 10.1111/2041-210X.12713).

There is also a non-CRAN version of a package for R available, `gruyere`

```{r}
## must have "cheddar" and "deSolve"

library(cheddar); library(deSolve)
devtools::install_github("quicklizard99/gruyere")
library(gruyere)

#library("gruyere", lib.loc="/Library/Frameworks/R.framework/Versions/3.5/Resources/library")
```

```{r}

```









