---
title: "Consumer-Resource Models with Variation"
output: github_document
---

One of the problems with the ATN model is that its goal is to generate "realistic" (or allometrically-constrained) networks of species that stably coexist. In modern coexistence theory, interested in how variation can influence coexistence.

## Example ODE in R

[Vignette](https://cran.r-project.org/web/packages/deSolve/vignettes/deSolve.pdf) for solving ODEs in R with package `deSolve`.

Load functions we'll use throughout

```{r, message=FALSE, warning=FALSE}
library(tidyverse); library(deSolve)
```

Example code for LV model in R:

```{r}
LotVmod <- function (Time, State, Pars) {
    with(as.list(c(State, Pars)), {
        dx = x*(alpha - beta*y)
        dy = -y*(gamma - delta*x)
        return(list(c(dx, dy)))
    })
}
 
Pars <- c(alpha = 2, beta = .5, gamma = .2, delta = .6)
State <- c(x = 10, y = 10)
Time <- seq(0, 100, by = 1)
 
out <- as.data.frame(ode(func = LotVmod, y = State, parms = Pars, times = Time))
 
matplot(out[,-1], type = "l", xlab = "time", ylab = "population")
```

## Vasseur & Fox 2007 Ecology Letters

Continuous time ODE of a 4 species system (resource, 2 consumers, 1 top predator), building on discrete time version in Ripa & Ives (2003, Theoretical Population Biology). Includes environmental fluctuation - though focus is more on the generation of population synchrony.

### V & F: no variation

Note that as written, without any environmental variation in consumer mortality, the dynamics follow those of McCann, Hastings, Huxel (1998, Nature)

```{r}
VassFox <- function (Time, State, Pars) {
    with(as.list(c(State, Pars)), {
      #M_C1_j = M_C1_0 * exp()
      #M_C2_j = M_C2_0 * exp()
      
      dP = - (M_P * P) + ( ( (J_P * P) * ( (O_P_C1 * C1) + ( (1 - O_P_C1) * C2) ) ) / ( (O_P_C1 * C1) + ((1 - O_P_C1) * C2) + C_0) ) 
      dC1 = - (M_C1_0 * C1) + ( (O_C1_R * J_C1 * C1 * R) / (R + R_0_1) ) - ( (O_P_C1 * J_P * P * C1) / ( (O_P_C1 * C1) + ((1 - O_P_C1) * C2) + C_0) )
      dC2 = - (M_C2_0 * C2) + ( (O_C2_R * J_C2 * C2 * R) / (R + R_0_2) ) - ( ((1 - O_P_C1) * J_P * P * C2) / ( ( O_P_C1 * C1) + ((1 - O_P_C1) * C2) + C_0) )
      dR = r * R * (1 - (R / K)) - ( (O_C1_R * J_C1 * C1 * R) / (R + R_0_1) ) - ( (O_C2_R * J_C2 * C2 * R) / (R + R_0_2) )
      
      return(list(c(dP, dC1, dC2, dR)))
      
    })
}


# from Table 1
pars <- c(
  # resource intrinsic rate of growth
  r = 1.0,
  # resource carrying capacity
  K = 1.0,
  # consumer 1 ingestion rate
  J_C1 = 0.8036,
  # consumer 2 ingestion rate
  J_C2 = 0.7,
  # predator ingestion rate
  J_P = 0.4,
  # medial consumer 1 mortality rate
  M_C1_0 = 0.4,
  # medial consumer 2 mortality rate
  M_C2_0 = 0.2,
  # predator mortality rate
  M_P = 0.08,
  # half saturation constant
  R_0_1 = 0.16129,
  R_0_2 = 0.9,
  C_0 = 0.5,
  # preference coefficient
  O_P_C1 = 0.92,
  O_C1_R = 1.0,
  O_C2_R = 0.98
)

State <- c(P = 1, C1 = 1, C2 = 1, R = 1)
Time <- seq(0, 1000, by = 1)
 
VF_out <- as.data.frame(ode(func = VassFox, y = State, parms = pars, times = Time))
plot.new()
matplot(VF_out[,-1], type = "l", xlab = "time", ylab = "population"); legend("topright", legend = c("P", "C1", "C2", "R"), col = 1:4, cex = 0.5, pch = 16)
```


### V & F: co-variation in consumer mortality

Will need to figure out how to implement the covariance (equations 1 & 2 in Vasseur & Fox, also equations 3 & 4 in Gouhier et al. 2010 Am Nat)

```{r}
VassFox_Cvar <- function (Time, State, Pars) {
    with(as.list(c(State, Pars)), {
      #M_C1_j = M_C1_0 * exp()
      #M_C2_j = M_C2_0 * exp()
      
      dP = - (M_P * P) + ( ( (J_P * P) * ( (O_P_C1 * C1) + ( (1 - O_P_C1) * C2) ) ) / ( (O_P_C1 * C1) + ((1 - O_P_C1) * C2) + C_0) ) 
      dC1 = - (M_C1 * C1) + ( (O_C1_R * J_C1 * C1 * R) / (R + R_0_1) ) - ( (O_P_C1 * J_P * P * C1) / ( (O_P_C1 * C1) + ((1 - O_P_C1) * C2) + C_0) )
      dC2 = - (M_C2 * C2) + ( (O_C2_R * J_C2 * C2 * R) / (R + R_0_2) ) - ( (O_P_C1 * J_P * P * C2) / ( (O_P_C1 * C1) + ((1 - O_P_C1) * C2) + C_0) )
      dR = r * R * (1 - (R / K)) - ( (O_C1_R * J_C1 * C1 * R) / (R + R_0_1) ) - ( (O_C2_R * J_C2 * C2 * R) / (R + R_0_2) )
      
      return(list(c(dP, dC1, dC2, dR)))
      
    })
}


sigma <- 0.55
ro <- 0.75

z <- matrix(data = rnorm(n = 2*1000, mean = 0, sd = 1), nrow = 2)
cholesky <- matrix(data = c(sigma^2, ro * (sigma ^2), ro * (sigma ^2), sigma ^2), 
                   nrow = 2)
g <- cholesky %*% z

eventfun <- function(t, y, M_C1_0, M_C2_0, g) {
  with (as.list(y),{
    M_C1 <- M_C1_0 * g[1,]
    M_C2 <- M_C2_0 * g[2,]
    return(c(M_C1, M_C2))
  })
}

# from Table 1
pars <- c(
  # resource intrinsic rate of growth
  r = 1.0,
  # resource carrying capacity
  K = 1.0,
  # consumer 1 ingestion rate
  J_C1 = 0.8036,
  # consumer 2 ingestion rate
  J_C2 = 0.7,
  # predator ingestion rate
  J_P = 0.4,
  # medial consumer 1 mortality rate
  M_C1_0 = 0.4,
  M_C1 = 0,
  # medial consumer 2 mortality rate
  M_C2_0 = 0.2,
  M_C2 = 0,
  # predator mortality rate
  M_P = 0.08,
  # half saturation constant
  R_0_1 = 0.16129,
  R_0_2 = 0.9,
  C_0 = 0.5,
  # preference coefficient
  O_P_C1 = 0.92,
  O_C1_R = 1.0,
  O_C2_R = 0.98
)

State <- c(P = 1, C1 = 1, C2 = 1, R = 1)
Time <- seq(0, 1000, by = 1)
eventtime <- seq(1, 1000, by = 1)
 
VF_out <- as.data.frame(ode(func = VassFox_Cvar, y = State, parms = pars, times = Time,
                            events = list(func = eventfun, 
                                          time = eventtime)))
```



