---
title: "Testing assumptions in the Forde & Doak model"
output: github_document
---

Quick plots and tests of various concepts that are embedded in the forde & doak model.

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
```

## fecundity & recruitment assumption

is local fecundity related to mussel recruitment density? forde & doak say no. test using data from sanford & menge 2007 meps.

```{r}
mussel_recruitment_density <- c(15.2, 27.2, 22.2, 19.4, 297.3, 115.1)
pyloric_caecum <- c(14.2, 13.5, 13.9, 12.2, 19.5, 11.3)
cor.test(mussel_recruitment_density, pyloric_caecum)
```

```{r, echo=FALSE}
plot(x = pyloric_caecum, y = mussel_recruitment_density)
```

no statistical relationship, just like f&d assumes


## potential recruitment assumptions

from forde & doak 2004 meps

potential recruitment function: should have saturating number of recruits as the number of larvae in the pool increases and free space decreases

```{r}
B.mean <- 50000
B.stdev <- sqrt(3.24*10^10)
location.B <- log(B.mean^2 / sqrt(B.stdev^2 + B.mean^2))
shape.B <- sqrt(log(1 + (B.stdev^2 / B.mean^2)))

a_x <- 3 * (10^-6) # barnacle recruit size from f&d
l_x <- rlnorm(n=100, location.B, shape.B)
F_tprev <- c(0.1, 0.5, 0.75, 0.9)
A_x <-  9.8 * (10^-5) # adult barn size

do_L_x <- function(a_x, l_x, F_tprev) (F_tprev / A_x) * (1 - exp(-(a_x * l_x )/ F_tprev))
```

```{r, echo=FALSE}
as_tibble(expand.grid(F_tprev, l_x, a_x)) %>%
  rename (F_tprev = "Var1", l_x = "Var2", a_x = "Var3") %>%
  mutate(L_x = do_L_x(a_x, l_x, F_tprev)) %>%
  mutate(F_tprev = factor(F_tprev)) %>%
  ggplot(aes(x = l_x, y = L_x, col = F_tprev)) +
  geom_point() + 
  geom_line(aes(col = F_tprev)) +
  labs(x = "number of larvae in the larval pool for species x (random Normal), time = t",
       y = "total number of potential recruits of species x, time = t",
       col = "free space, time = t-1")
```

behaves the way we think it should


## actual recruitment assumptions

should never get more actual recruits than potential recruits..

```{r}
# given potential recruitment numbers for two species
L_x <- c(100, 1000, 2500, 5000, 10000)
L_y <- c(100, 1000, 2500, 5000, 10000)
# similar adult sizes
A_x <- 5 * (10^-5)
A_y <- A_x
# and a range of free space
F_tprev <- c(0.1, 0.5, 0.75, 0.9)

do_R_x <- function(F_tprev, L_x, A_x, L_y, A_y) (F_tprev / ((L_x * A_x) + (L_y * A_y))) * L_x
```

```{r, echo = FALSE}
as_tibble(expand.grid(F_tprev, L_x, A_x, L_y, A_y)) %>%
  rename (F_tprev = "Var1", L_x = "Var2", A_x = "Var3", 
          L_y = "Var4", A_y = "Var5") %>%
  mutate(R_x = do_R_x(F_tprev, L_x, A_x, L_y, A_y)) %>%
  mutate(F_tprev = factor(F_tprev),
         L_y = factor(L_y)) %>%
  ggplot(aes(x = L_x, y = R_x, col = L_y)) +
  geom_point() + 
  geom_line(aes(col = L_y)) + 
  facet_wrap (~ F_tprev) +
  geom_abline(slope = 1, intercept = 0) +
  labs(y = "number of actual recuits of species x",
       x = "number of potential recruits of species x",
       col = "number of potential\nrecruits of competitor species")
```

often see that the number of actual recruits >> potential recruits in the larval pool, as a function of multiple factors. this is likely due to the form of the function.

instead rewrite function following iwasa & roughgarden 1986

```{r}
# here, the abundance of larvae of the ith species is equation 3 in I & R, but just assumed to be a random lognormal draw in F & D (l_x)
do_F_prev <- function(A, a_x, X, a_y, Y) {A - ((a_x * X) + (a_y * Y))}
do_R_x_ir <- function(c_x, L_x, F_tprev) c_x * L_x * F_tprev

A <- 1
X_0 <- c(1, 1000, 10000, 100000)
Y_0 <- c(1, 1000, 10000, 100000)
a_x <- 3 * (10^-6)
a_y <- 3 * (10^-6)
# recruit supply rate c should = sr, which is 0.7 in f&d
c_x <- c()
L_x <- 1000



expand.grid(A=A, a_x=a_x, a_y=a_y, c_x=c_x, L_x=L_x, X_0=X_0, Y_0=Y_0) %>%
  as_tibble() %>%
  mutate(F_tprev = do_F_prev(A, a_x, X_0, a_y, Y_0)) %>%
  mutate(R_t = do_R_x_ir(c_x, L_x, F_tprev))



```

